---
title: "Sensitivitty_rew_startup - correlation and data cleaning"
author: "Liv Tollånes"
date: "2023-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load("tidyverse", "ggpubr", "psych", "Hmisc", "corrplot", "rstatix")

```


## Task Nr.1 - Avergaing Scores across POS/NEG CBM training performance
- Average participants' performance across the 2 EEfRt tasks to create three averaged scores: one for low, one for medium, and one for high probability trials). 
- Then correlate with baseline measures.

### Data loading and format alterations
```{r}
#Loading in data
vars_CBM <- read.csv("./data/CBM_variables.csv")
df_CBM <- read.csv("./data/CBM_data.csv")


#Creating subset for just for better overview
#EEfRT_sub <- df_CBM %>% select(c((1:7), contains("EEfRT")))

length(unique(df_CBM $participant_id)) #81 unique subs

# Select only IDs with exclude values 0. 2 rows were dropped - so now there are 79 participants in the df
df_CBM  <- subset(df_CBM ,exclude == 0 )

#Replace weird column values with NA
df_CBM [df_CBM  == "#NULL!"] <- NA

#Remove NAs before Creating the averaged scores

as_tibble(df_CBM )


# Make all EEfRT columns numeric
# MAKE THIS CODE LESS MANUAL AT SOME POINT
# Pos high prob
df_CBM $POS_EEfRT_prop_hard_choice_highprob <- as.numeric(df_CBM $POS_EEfRT_prop_hard_choice_highprob, na.rm = T)
is.numeric(df_CBM $POS_EEfRT_prop_hard_choice_highprob)

#Pos med prob
df_CBM $POS_EEfRT_prop_hard_choice_medprob <- as.numeric(df_CBM $POS_EEfRT_prop_hard_choice_medprob, na.rm = T)
is.numeric(df_CBM $POS_EEfRT_prop_hard_choice_medprob)

#POS low prob
df_CBM $POS_EEfRT_prop_hard_choice_lowprob <- as.numeric(df_CBM $POS_EEfRT_prop_hard_choice_lowprob, na.rm = T)
is.numeric(df_CBM $POS_EEfRT_prop_hard_choice_lowprob)

#NEG high prob
df_CBM $NEG_EEfRT_prop_hard_choice_highprob <- as.numeric(df_CBM $NEG_EEfRT_prop_hard_choice_highprob, na.rm = T)
is.numeric(df_CBM $NEG_EEfRT_prop_hard_choice_highprob)

#NEG med prob
df_CBM $NEG_EEfRT_prop_hard_choice_medprob <- as.numeric(df_CBM $NEG_EEfRT_prop_hard_choice_medprob, na.rm = T)
is.numeric(df_CBM $NEG_EEfRT_prop_hard_choice_medprob)

#NEG low prob
df_CBM $NEG_EEfRT_prop_hard_choice_lowprob <- as.numeric(df_CBM $NEG_EEfRT_prop_hard_choice_lowprob, na.rm = T)
is.numeric(df_CBM $NEG_EEfRT_prop_hard_choice_lowprob)

#Create Average column (average of positive and negative EEfRT scores averaged together into one EEfRT total score)
df_CBM$EEfRT_averaged_lowp <- rowMeans(df_CBM [ ,c("POS_EEfRT_prop_hard_choice_lowprob", "NEG_EEfRT_prop_hard_choice_lowprob")])
df_CBM$EEfRT_averaged_medp <- rowMeans(df_CBM [ ,c("POS_EEfRT_prop_hard_choice_medprob", "NEG_EEfRT_prop_hard_choice_medprob")])
df_CBM$EEfRT_averaged_highp <- rowMeans(df_CBM [ ,c("POS_EEfRT_prop_hard_choice_highprob", "NEG_EEfRT_prop_hard_choice_highprob")])

#Saving the data in a finished preprocessing file - to not repeat the above code every time
write.csv(df_CBM, "./data/preprocessed.csv", row.names=FALSE)

```


```{r}
# Loading in preprocessed data - when running code again
 df <- read.csv("./data/preprocessed.csv")
```

### Correlate with baseline measures 
- Are there any relationships between the baseline measures and the averaged EEfRT scores?
- Test strength of association between variables

```{r}
# Subset all columns containing EEfRT and the baseline measures - to make the task of correlations easier to overview
# Baseline measures include all MASQ-, TEPS, and BIS/BAs-columns
sub <- df %>% select(c((1:7), EEfRT_averaged_lowp, EEfRT_averaged_medp, EEfRT_averaged_highp, TEPS_ANT, TEPS_CON, MASQ_GD, MASQ_AD, MASQ_AA, BAS_Drive, BAS_FS, BAS_RR, BIS))


# #Removing non-useful columns in order to make the corr matrix work easily
sub$exclude <- NULL 
sub$participant_id <- NULL
# sub$Nationality <- NULL
# sub$Gender <- NULL

#Testing the class of all variables at once
sapply(sub, class)

#Change all baseline measures to numeric
sub <- sub %>% mutate_at(c(6:17), as.numeric) #Remember to update the number specified according to the number of columns dropped

# na.omit(sub)  shuld NAs be omitted for the correlation calculation?


# Asummption of normally distributed data for pearson's correlation test
shapiro.test(sub$EEfRT_averaged_lowp) # p<.05 - not normal
shapiro.test(sub$EEfRT_averaged_highp) # p <.05 not normal
shapiro.test(sub$EEfRT_averaged_medp) #p>.05 - normal


hist(sub$EEfRT_averaged_lowp, col = 'coral2')
hist(sub$EEfRT_averaged_highp, col = 'steelblue')
hist(sub$EEfRT_averaged_medp)

```

```{r}
#### Correlation Work - IF TIME: create the correlation plot with their corresponding p-values

#creating a subset for correlation work
corsub <- sub %>% select(c(, (6:17)))
corsub <- na.omit(corsub)
sapply(corsub, class)
corsub_matrix <- as.matrix(corsub)


#create Spearman correlation matrix
corr_mat <-  cor(corsub, method="spearman")

#Create a p-value matrix for including in the plot
#cor_matrix <- cor_pmat(sub, method = "spearman")

#Correlation matrix - using nonparametric Spearman correlation test due to non-normal variables, and not a complete set of continuous variables
#Obtains p-values as well 
rcorr(corsub_matrix, type = c("spearman"))

#plotting the correlation matrix
corrplot(corr_mat, method = c("number"), type = "upper") #With specific numbers
corrplot(corr_mat) #With circles to represent correlations instead

```

Output of the correlation analysis
- For the high and medium probability categories, no relationship between the averaged EEfRT scores and baseline measures were found (p>.05)
- For the low probability category, a small positive correlation (0.29) was found between the averaged EEfRT score and the MASQ.AA measure (p<0.05). Otherwise, no significant correlations were found for this probability category and the rest of the measure
- Regarding the predictor variables - there seems to be a small positive correlation between the low and high probability EEfRT scores (p<0.05)


```{r}
#Descriptive stats - describing the data sample
sapply(sub, class)
sub <- na.omit(sub)

#Number of participants in each gender group
sub %>% count(Gender, sort = TRUE) #45 women, 18 men

#Number of participants in each nationality group
sub %>% count(Nationality, sort = TRUE) #There are 24 different nationalities, with most people coming from Britain



```






