---
title: "Models"
author: "Liv Tollånes"
date: "2023-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 2. Multilevel modelling of reward anticipation and self reports of pleasure and experience of reward
Q: Does lab-measured reward anticipation (averaged EEfRT scores) predict self-reported (i) "anticipated reward pleasure" (ii) "experienced reward" in the real world (ESM)? 


	- Use the averaged EEfRT scores created in task 1. (Scores from the 2 days(pos and neg) merged together, but with 3 levels, low, med, high probability)
	
	- Use multilevel modelling, with 3 levels of nested data (questionnaires, nested within days, within participants)
	
	- The analysis would be very close to the paper: https://psyarxiv.com/fnhd9/, but with EEfRT performance as the participant-level predictors (low, med, high) instead of the ISI, RPA scores, and ERQ scores. The EEfRT are obviously not day-level predictors, but more like our covariates baseline, ISI  - instead of baseline participant-level ISI scores, RPA scores, and ERQ scores). "Participant" and "Days within participant" are random intercepts.


```{r}
#Loading pa
```

