---
title: "Models"
author: "Liv Tollånes"
date: "2023-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 2. Multilevel modelling of reward anticipation and self reports of pleasure and experience of reward
Q: Does lab-measured reward anticipation (averaged EEfRT scores) predict self-reported (i) "anticipated reward pleasure" (ii) "experienced reward" in the real world (ESM)? 


	- Use the averaged EEfRT scores created in task 1. (Scores from the 2 days(pos and neg) merged together, but with 3 levels, low, med, high probability)
	
	- Use multilevel modelling, with 3 levels of nested data (questionnaires, nested within days, within participants)
	
	- The analysis would be very close to the paper: https://psyarxiv.com/fnhd9/, but with EEfRT performance as the participant-level predictors (low, med, high) instead of the ISI, RPA scores, and ERQ scores. The EEfRT are obviously not day-level predictors, but more like our covariates baseline, ISI  - instead of baseline participant-level ISI scores, RPA scores, and ERQ scores). "Participant" and "Days within participant" are random intercepts.


```{r}
#Loading packages 

pacman::p_load("tidyverse", "ggpubr", "Hmisc", "corrplot", "rstatix", "psych", "nlme", "lme4", "lmerTest", "JWileymisc","multilevelTools", "texreg")


```

```{r}
#Loading ESM data

ESM <- read.csv("./data/ESM_reduced.csv")
ESM_keys <- read.csv("./data/ESM_keys.csv")

unique(ESM$Actual_ID) # P022 was never there from the beginning

### Removing the excluded rows for low participation

#Are there still 2-s in the column?
ESM$Excluded.for.Low.Participation..1...no..2...yes. <-  as.factor(ESM$Excluded.for.Low.Participation..1...no..2...yes.)
is.factor(ESM$Excluded.for.Low.Participation..1...no..2...yes.)
unique(ESM$Excluded.for.Low.Participation..1...no..2...yes.)

# keeping an overview of what participants were removed - for matching with the CBM df
ESM_exluded_low_part <- filter(ESM, Excluded.for.Low.Participation..1...no..2...yes. == "2")
ESM_exluded_low_part_ID <- unique(ESM_exluded_low_part$Actual_ID) #The IDs removed are P010, P019, P020, P028, P038, P041, P049, P506, P513

#Removing from main df
ESM <- filter(ESM, Excluded.for.Low.Participation..1...no..2...yes. != "2")
unique(ESM$Excluded.for.Low.Participation..1...no..2...yes.)

length(unique(ESM$Actual_ID)) #There are 67 unique participants in this df



##############################
#CBM data set
CBM <- read.csv("./data/preprocessed.csv")
CBM_vars <- read.csv("./data/CBM_variables.csv")

length(unique(CBM$ID)) #There are 79 unique IDs - 12 more than in the ESM data

#Replace weird column values with NA
CBM[CBM  == "#NULL!"] <- NA


#Selecting only the relevant columns in the CBM data for better overview
CBM_sub <- CBM %>% select(c(1:7), 
                          EEfRT_averaged_lowp, 
                          EEfRT_averaged_medp, 
                          EEfRT_averaged_highp,
                          EEfRT_LowProb_DiffScore,
                          EEfRT_MedProb_DiffScore,
                          EEfRT_HighProb_DiffScore,
                          POSCBMTrainingaccuracy,	
                          POSCBMTestingaccuracy,
                          NEGCBMTrainingaccuracy,
                          NEGCBMTestingaccuracy,
                          RT_POS_TRAINING_IDX,
                          RT_NEG_TRAINING_IDX,
                          RT_Diff_Score,
                          TEPS_ANT, TEPS_CON, MASQ_AA, MASQ_GD, MASQ_AD,  BAS_Drive, BAS_FS,BAS_RR, BIS)

CBM_sub$exclude <- NULL
CBM_sub$participant_id <- NULL

# 16 participants have NAs in their EEfRT columns - due to lacking EEfRT scores

#Removing the participants not included in the ESM df first to inspect what IDs are left
CBM_sub <- filter(CBM_sub, ID != "P010")
CBM_sub <- filter(CBM_sub, ID != "P019")
CBM_sub <- filter(CBM_sub, ID != "P020")
CBM_sub <- filter(CBM_sub, ID != "P028")
CBM_sub <- filter(CBM_sub, ID != "P038")
CBM_sub <- filter(CBM_sub, ID != "P041")
CBM_sub <- filter(CBM_sub, ID != "P049")
CBM_sub <- filter(CBM_sub, ID != "P506")
CBM_sub <- filter(CBM_sub, ID != "P513")

length(unique(CBM_sub$ID)) #Length is now 70, but 12 problematic participants are left. Only 4 of those removed were NA-participants

# Looking into the original df for these participants - what is the issue?
#The problematic IDs are: P003, P024, P036, P501, P509, P510, P512, P515, P525, P526, P535, P536
probID <- subset(CBM, ID %in% c("P003", "P024", "P036", "P501", "P509", "P510", "P512", "P515", "P525", "P526", "P535", "P536"))


#Further inspection shows that the problematic IDs are caused by NAs in the various prob categories of the EEfRT measurements - so incomplete CBM data. Removing all of them
CBM_sub <- CBM_sub %>% filter(!ID %in% c("P003", "P024", "P036", "P501", "P509", "P510","P512", "P515", "P525", "P526", "P535", "P536")) #The resulting df 


#Unique IDs in the subsetted CBM data
length(unique(CBM_sub$ID)) #The resulting df now has 58 unique participants. There were 67 in the ESM data. 
CBM_sub_unique_IDs <- unique(CBM_sub$ID)

#Looking at which IDS in the ESM df do not occur in the CBM df
unique(ESM$Actual_ID)

unique(CBM_sub$ID)

#The IDs that occur in the ESM data but not in the CBM data are:
#P024, P036, P501, P509, P512, P525, P526, P535, P536 (9 in total)
#Removing them from the ESM data:
ESM <- ESM %>% filter(!Actual_ID %in% c("P003","P024", "P036", "P501", "P509", "P512", "P525", "P526", "P535", "P536"))

unique(ESM$Actual_ID)

unique(CBM_sub$ID)

#P022 is in CBM sub - but was never there for ESM. Removing P022 from CBM_sub
CBM_sub <- CBM_sub %>% filter(!ID %in% "P022")

length(unique(ESM$Actual_ID))
length(unique(CBM_sub$ID)) #The two dfs are now equal 




##### Instead of all of the above, I guess an alternative option was to just merge the dfs, and write na.action "na-omit" in the model .....
```



```{r}
#The next task is to add three EEfRT columns to the ESM data frame, replicating their scores across each row every partipant has
#Rename Actual_ID in ESM df to ID
ESM <- ESM %>% rename(ID = Actual_ID)

#Creating a merged column of subject and day
ESM$subj_day <- paste(ESM$ID, ESM$Day)

#Merging the two dfs
ESM_merged <- merge(ESM, CBM_sub, all.x=TRUE, by="ID")
ESM_merged <- as_tibble(ESM_merged)

#reordering columns for better overview
ESM_merged <- ESM_merged  %>% relocate(subj_day, .after=Day)
ESM_merged <- ESM_merged  %>% relocate(Excluded.for.Low.Participation..1...no..2...yes., .after=BIS)
ESM_merged <- ESM_merged  %>% relocate(StartDate, .after=Excluded.for.Low.Participation..1...no..2...yes.)
ESM_merged <- ESM_merged  %>% relocate(Nationality, .after=StartDate)


```

## Actual modelling

Variable types:
- Outcome/dependent variables (ant/con. pleasure) are ordinal (measured on 7 pt. Likert scale)
- Predictors: EEfRT averages are ratios. Between 0 and 1


I should consider centering the outcome variables (ant/con pleasure - see link: https://philippmasur.de/2018/05/23/how-to-center-in-multilevel-models/)

Christine writes that the multilevel analysis should contain:
- checking residuals
- mean-centring some variables
- running models (with and without outliers)
- generating some plots at the end

```{r}
#Inspecting range of EEfRT ratios (predictors)
min(ESM_merged$EEfRT_averaged_lowp) #0
max(ESM_merged$EEfRT_averaged_lowp) #0.81

min(ESM_merged$EEfRT_averaged_medp) #0
max(ESM_merged$EEfRT_averaged_medp) #0.975

min(ESM_merged$EEfRT_averaged_highp) #0.03
max(ESM_merged$EEfRT_averaged_highp) #0.1


#Consider centering the outcome variable
```




```{r}
ESM %>%
    group_by(Actual_ID) %>%
    summarise(count=n())
```












